Variability of Planetary Mass Companion 2M1207 B

### 2015-08-29 summary/letter/letter_vAC.tex
new bugs, never stop

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shorttitle{Short Title}
\shortauthors{Zhou et al.}

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\newcommand{\tinytim}{\textit{Tiny Tim}}
\begin{document}
\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider, ...}
\affil{The University of Arizona}

@@

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist
sampled PSF, but significant systematic errors of \tinytim{} PSF
 limits its ability in high precision
photometry\citep{Biretta2014}. However, 6 orbits of observation provides us a chance to
fully characterize the difference of model and observed PSFs. To obtain
robust \tinytim{} PSF photometry,  we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to better model the cold mask,
diffraction spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
We determined the exact position of 2M1207A first by fitting the PSF
on a dynamic coordinate grid and minimizing the difference in the
observed and modeled PSFs at a region centered on 2M1207 A with a
5-pixel-radius aperture centered 2M1207 b excluded.Then we fit the
position of 2M1207 b and the photometries of 2M1207 A and b together
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs are very stable for
given PSF position. Therefore at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
empirical correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{model} - PSF_{obs} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs, and fitted for their
amplitudes together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Because the total fluxes of the model PSFs were
normalized to unity as default, the fluxes of 2M1207 A and B were
solely represented by the amplitude of the two PSFs coordinately.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
  verification tests. Individual measurements are plotted with gray
  crosses. Photometries of the same exposure sequence are binned, and
  binned data are plotted with points and squares. Fitted sinusoidal
  waves are plotted with solid lines. {\em Upper}: binned photometries taken
in dithering position 1 and 3 (red points) and that taken in 2 and 4
(blue squares) are plotted differently. They demonstrate same trend of
variation. In upper left panel, sinusoidal wave fitted with all
parameter set free, and that fitted with period set the same as that
of F125W are plotted with green and purple lines, respectively. {\em
  Middle}: sinusoidal waves fitted without using the data taken in
Orbit 1 are plotted. These curves are almost identical to the curves
plotted in upper panel. {\em Lower}: photometry measured with
AFEM-applied images and their best fitted sinusoidal curves are
plotted. These photometries and curves are also almost identical to
those plotted in upper panel.}
  \label{fig:2}
\end{figure*}

@@

To constrain the uncertainties of the least square optimization
results, we used a Monte Carlo(MC) method to improve the fitting. We
generated a series of random Gaussian noises with the standard deviation same as
the photon noise, added them to the original light curves, and applied
least square fit to the new light curves. We repeated above routine for 10000 times and obtained
the distribution of the fitting parameters. The distributions for the
periods and the amplitudes for F125W and F160W light curves are
shown in Figure \ref{fig:4}.

@@

\begin{figure*}
  \centering
  \plottwo{periodDistr}{amplitudeDistr}
  \caption{Distributions for periods (left) and amplitudes(right) for the light
    curve of F125W and F160W. The bin size for histograms of period is
  0.25 hour and for that of amplitude is 0.5\%. Histograms are
  normalized in the way that total area of the histogram equals to
  1. In the right panel, Gaussian profiles are fitted to the
  histograms and plotted in solid lines.}
  \label{fig:4}
\end{figure*}

\label{Results}
We present the first high-contrast, high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the 5--7~M$_{J}$ companion 2M1207b, the first detection
of modulations in directly imaged ultracool objects.  The best fit
periods for F125W and F160W are 10.4 and 9.0 hour correspondingly. The
amplitudes for the normalized light curves are 1.4\% and 0.9\% for
F125W and F160W light curves.

@@

In the following we will discuss the amplitude and period of 2M1207B
and place this object in the broader context of ultracool atmospheres.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With probability range
of 64\%(1-$\sigma$ Gaussian), we estimated the 1-$\sigma$ range for
the periods of F125W and F160W to be  $10.5_{-1.2}^{+1.3}$ and
$9.1_{-1.0}^{+1.1}$ h, respectively. The period for best
fitted sine wave of F125W light curve is longer than that of F160W for
1.5 h that is slight larger 1-$\sigma$ standard deviation. We also
performed a joint MC fit with the two light curves with the assumption
that the periods of the variation of the two bands are the same. We
derived a period of $10.2^{+0.9}_{-0.8}$ h.

The variation amplitudes in the two bands have significant
difference. The distributions of the amplitudes are well described by
Gaussian profiles. Fitting the histogram to a Gaussian function, we
determined that amplitude distribution of F125W peaks at 1.5\% with a
standard deviation of 0.3\%, and that of F160W have mean and standard
deviation of 0.9\% and 0.2\%, respectively. The peaks of the two
histograms separated by more than 2-$\sigma$. The variation amplitude
of F125W light curve is 1.67 times of that of F160W light curve.

@@

\section{Discussion}
Amplitude and Relative Amplitudes: Comparison to other brown dwarfs

Rotation Period: Comparison to Brown Dwarfs and Beta Pic

Method: Opens a new window, successful demonstration, how could it be improved?
Mapping different pressure levels, atmospheric dynamics?

\section{Conclusions}
\begin{itemize}
\item a data-reduction pipeline is developed to obtain high precision
  photometry measurement from high contrast WFC3 IR data. For a
  contrast of $\sim 7$ magnitudes at an angular separation of
  $\sim0.7''$, we obtained photometry measurement for 2M1207 B at
  precision of about 2-3\%. 
\item time variability for 2M1207 B was discovered, the light curve of
  2M1207 B for both two colors can be fitted with a T=10.7 hr
  sinusoidal curve.
\item constrain on the amplitude of the amplitude of the
  variability. Large amplitude can be excluded. Constraints on the
  inhomogeniety of the cloud coverage can be inferred -- small
  thickness variance...
\item ? The atmosphere and cloud structure of 2M1207 B. How does it
  compare to the brown dwarf.
\end{itemize}

>>>>

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{apj}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\shortauthors{Zhou et al.}

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\newcommand{\tinytim}{\textit{Tiny Tim}}
\newcommand{\bpic}{$\beta$ Pic}
\newcommand{\vsini}{$v\sin i$}
\newcommand{\mjup}{M$_{\mbox{Jup}}$}
\begin{document}

\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider, ...}
\affil{The University of Arizona}

@@

One major challenge of high contrast imaging observation using WFC3/IR
is significant under-sampling of the detector.  2M1207 A and b are
only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF. When
applying roll subtraction, notable artifact structures are generated
by image interpolation and shifting. On the other hand, \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist
sampled PSF, but significant systematic errors of \tinytim{} PSF
 limits its ability in high precision
photometry\citep{Biretta2014}. However, 6 orbits of observation
provides us an opportunity to
fully characterize the difference of model and observed PSFs. To obtain
robust \tinytim{} PSF photometry,  we designed a 2-round PSF fitting strategy: 1. calculating
correction map for \tinytim{}; 2. hybrid PSF photometry.

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to better model the cold mask,
diffraction spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
We determined the exact position of 2M1207A first by fitting the PSF
on a dynamic coordinate grid and minimizing the difference in the
observed and modeled PSFs at a region centered on 2M1207 A with a
5-pixel-radius aperture centered on 2M1207 b excluded. Then introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b together
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs are very stable for
given PSF position. Therefore at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
empirical correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{model} - PSF_{obs} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and scaled the three component together to get best fit. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Because the total fluxes of the model PSFs were
normalized to unity as default, the fluxes of 2M1207 A and B were
solely represented by the amplitude of the two PSFs.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
  variability verification tests. Individual measurements are plotted with gray
  crosses. Photometries of the same exposure sequence are binned, and
  binned data are plotted with points or squares. Fitted sinusoidal
  waves are plotted with solid lines. {\em Upper}: binned photometries taken
in dithering position 1 and 3 (red points) and that taken in 2 and 4
(blue squares) are plotted differently. They demonstrate same trend of
variation. In upper left panel, sinusoidal wave fitted with all
parameter set free, and that fitted with period set the same as that
of F125W are plotted with green and purple lines, respectively. {\em
  Middle}: sinusoidal waves fitted without using the data taken in
Orbit 1 are plotted. These curves are almost identical to the curves
plotted in upper panel. {\em Lower}: photometry measured with
AFEM-applied images and their best fitted sinusoidal curves are
plotted. These photometries and curves are also almost identical to
those plotted in upper panel.}
  \label{fig:2}
\end{figure*}

@@

To constrain the uncertainties of the least square optimization
results, we used a MC method to improve the fitting. We
generated a series of random Gaussian noises with the standard deviation same as
the photon noise, added them to the original light curves, and applied
least square fit to the new light curves. We repeated above routine for 100,000 times and obtained
the distribution of the fitting parameters. The distributions for the
periods and the amplitudes for F125W and F160W light curves are
shown in Figure \ref{fig:4}.

\begin{figure*}
  \centering
  \includegraphics[width=0.45\textwidth]{periodDistr}
  \includegraphics[width=0.45\textwidth]{amplitudeDistr}
  \caption{Distributions for periods (left) and amplitudes(right) for the light
    curve of F125W and F160W. The bin size for histograms of period is
  0.25 hour and for that of amplitude is 0.5\%. Histograms are
  normalized in the way that total area of the histogram equals to
  1. In the right panel, Gaussian profiles are fitted to the
  histograms of periods and plotted in solid lines.}
  \label{fig:4}
\end{figure*}

@@

\label{Results}
We present the first high-contrast, high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the 5--7~M$_{J}$ companion 2M1207b, the first detection
of modulations in directly imaged ultracool objects.  The best fit
periods for F125W and F160W are 10.5 and 9.1 hour correspondingly. The
amplitudes for the normalized light curves are 1.45\% and 0.92\% for
F125W and F160W light curves.

@@

% In the following we will discuss the amplitude and period of 2M1207B
% and place this object in the broader context of ultracool atmospheres.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimated the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sine wave of F125W light
curve is 1.5h longer than that of F160W  that is slight larger
1-$\sigma$ standard deviation. We also jointly fit the two band light
curve together forcing the periods of two sine waves to be the
same. We derived a modulation period of $10.2^{+0.9}_{-0.8}$ h in this
circumstance.

We discovered that  the variation amplitudes in the two bands were significantly
different. The distributions of the amplitudes are well described by
Gaussian profiles. By fitting a Gaussian function to the distribution, we
determined that amplitude distribution of F125W peaks at 1.45\% with a
standard deviation of 0.22\%, and that of F160W have mean and standard
deviation of 0.92\% and 0.20\%, respectively. The peaks of the two
histograms separated by more than 2-$\sigma$. The variation amplitude
of F125W light curve is 1.58 times of that of F160W light curve.

@@

\section{Discussion}

\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change with brown
    dwarfs, \bpic{} b, and solar system planets. {\em Left}: period vs. mass
  plot for 2M1207b (red square), solar system planets and \bpic{} b
  (blue squares), and brown dwarfs (black circles, gray shade). The mass of brown
  dwarfs are assumed to be 30 \mjup{}. The gray rectangle
  ranging from 15 \mjup{} to 50 \mjup{} in $x$, and $\pm\sigma$
  with the center at mean rotation period of brown dwarfs in $y$, indicates a
region where brown dwarfs most likely to appear in this diagram. {\em Right}: ratio
of variation amplitude in J and H band vs. spectral type for 2M1207b
and brown dwarfs. Points are
color coded with their J$-$H colors. }
  \label{fig:5}
\end{figure*}

A fundamental result of our study is the direct determination of the
rotation period of a directly imaged planetary-mass object. In
Figure~\ref{fig:5} we compare the rotation period of 2M1207b to the
solar system planets, field brown dwarfs that have well defined
rotation period from the study of
\citep[][]{Metchev2015}, and \bpic{} b the only other directly imaged
planet with an estimated period from measured \vsini{}.  The study by
\citep[][]{Snellen2014} succeeded in mesuring \vsini{} for \bpic{} b
and demonstrated that it fits a trend defined by Solar System planets
in which more massive planets have faster rotation rates. The
interesting finding that \bpic{} b, an exoplanet that formed in a
protoplanetary disk, follows this trend suggests a possibly connection
between planet mass, initial angular momentum, and formation in a
disk.

Excitingly, our measurement of the rotation period of 2M1207b, an
object with similar mass and age to \bpic{} b, show that their rotation
periods are very similar, {\em in spite } of the fact that 2M1207b 
most likely not formed in a circumstellar (or circumsubstellar) disk
but is a result of gravitational fragmentation. This suggests
that rotation periods are not good tracers of the formation pathways
and may not contribute important evidence for a formation in a disk
vs. in a cloud core environment.

Our observations, furthermore, allow us to compare the relative
amplitudes in the J- and H-bands between the handful of brown dwarfs
for which high-quality near-infrared time-resolved observations have
been obtained. In the right panel of Figure \ref{fig:5}, we plot the
relative amplitude of J- and H-bands of 2M1207b with brown dwarfs
\citep{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014} that
have different spectral types and J$-$H colors.

We found a strong correlation between the spectral type of the object
and the J to H variation amplitude ratio. Figure \ref{fig:5}
demonstrates that earlier spectral type objects -- independent of
their surface gravity -- have larger variations at shorter wavelength
than at longer wavelengths. Importantly, although the J$-$H color of
2M1207b is significantly redder the other L5 dwarfs, we note that its
relative variation amplitude ratio is almost identical to the matching
spectral type mid-L dwarf 2M1821. This suggests that the upper cloud
layers, probed through variability, may be more influenced by spectral
type and less by cloud thickness, which is thought to influence the
J$-$H color.

{\bf Two questions: What processes sets the relative J vs H amplitude
  and why is it tied to the pectral type and not to the NIR colors or
  surface gravity? What does this tell us about the cloud structure
  as a function of spectral type / gravity (even if speculation)?}

\section{Conclusions}
In summary from our J- and H-band high precision,
high-cadence lightcurves we discovered sinusoidal modulations in the  planetary mass object
2M1207b. This is the first detection of rotational modulations in a
directly imaged planetary mass object.  
The period is  $10.2^{+0.9}_{-0.8}$, very similar to that derived from
{\em v sin i} measurements for  the direct imaged exoplanet \bpic{} b and significantly longer than
most field brown dwarfs with known rotation periods. The
relative modulation amplitude of J and H band is almost identical to
one matching spectral type L5 dwarf, although they have very different
J$-$H colors, and it is markedly different form later spectral type
brown dwarfs.

Finally, we note that the observations presented here open an exciting
new window on directly imaged exoplanets and planetary-mass
objects. Our study demonstrates a successful application of high-cadence,
high-precision, high-contrast photometry with planetary mass
companion. We also show that these observations can be carried out simultaneously
at multiple wavelengths, allowing us to prove multiple pressure
levels. With observation of a larger sample and at multiple wavelengths, we will be
able to explore  the detailed structures of atmospheres of directly
imaged exoplanets, and identify the key parameters that determine these.
% % \begin{itemize}
% % \item a data-reduction pipeline is developed to obtain high precision
% %   photometry measurement from high contrast WFC3 IR data. For a
% %   contrast of $\sim 7$ magnitudes at an angular separation of
% %   $\sim0.7''$, we obtained photometry measurement for 2M1207 B at
% %   precision of about 2-3\%. 
% % \item time variability for 2M1207 B was discovered, the light curve of
% %   2M1207 B for both two colors can be fitted with a T=10.7 hr
% %   sinusoidal curve.
% % \item constrain on the amplitude of the amplitude of the
% %   variability. Large amplitude can be excluded. Constraints on the
% %   inhomogeniety of the cloud coverage can be inferred -- small
% %   thickness variance...
% % \item ? The atmosphere and cloud structure of 2M1207 B. How does it
% %   compare to the brown dwarf.
% \end{itemize}
